### 2021-02-02 book/sources/BBhoare.tex
Latest

\>{\tiny 1}\'\>\texttt{\textbf{import}}~hoare\\

\>{\tiny 2}\'\>\\

\>{\tiny 3}\'\>\texttt{\textbf{def}}~BB(\textit{size}):\\

\>{\tiny 4}\'\>~~~~\textit{result}~=~\{\\

\>{\tiny 5}\'\>~~~~~~~~~~~~.mon:~hoare.Monitor(),\\

\>{\tiny 6}\'\>~~~~~~~~~~~~.prod:~hoare.Condition(),~.cons:~hoare.Condition(),\\

\>{\tiny 7}\'\>~~~~~~~~~~~~.buf:~\{~\textit{x}:()~\texttt{\textbf{for}}~\textit{x}~\texttt{\textbf{in}}~\{1..\textit{size}\}~\},\\

\>{\tiny 8}\'\>~~~~~~~~~~~~.head:~1,~.tail:~1,\\

\>{\tiny 9}\'\>~~~~~~~~~~~~.count:~0,~.size:~size\\

\>{\tiny 10}\'\>~~~~~~~~\}\\

\>{\tiny 11}\'\>~~~~\\

\>{\tiny 12}\'\>\texttt{\textbf{def}}~put(\textit{bb},~\textit{item}):\\

\>{\tiny 13}\'\>~~~~hoare.enter(?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 14}\'\>~~~~\texttt{\textbf{if}}~\textit{bb}$\rightarrow$count~==~\textit{bb}$\rightarrow$size:\\

\>{\tiny 15}\'\>~~~~~~~~hoare.wait(?\textit{bb}$\rightarrow$prod,~?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 16}\'\>~~~~\textit{bb}$\rightarrow$buf[\textit{bb}$\rightarrow$tail]~=~\textit{item}\\

\>{\tiny 17}\'\>~~~~\textit{bb}$\rightarrow$tail~=~(\textit{bb}$\rightarrow$tail~\%~\textit{bb}$\rightarrow$size)~+~1\\

\>{\tiny 18}\'\>~~~~\textit{bb}$\rightarrow$count~+=~1\\

\>{\tiny 19}\'\>~~~~hoare.signal(?\textit{bb}$\rightarrow$cons,~?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 20}\'\>~~~~hoare.exit(?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 21}\'\>\\

\>{\tiny 22}\'\>\texttt{\textbf{def}}~get(\textit{bb}):\\

\>{\tiny 23}\'\>~~~~hoare.enter(?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 24}\'\>~~~~\texttt{\textbf{if}}~\textit{bb}$\rightarrow$count~==~0:\\

\>{\tiny 25}\'\>~~~~~~~~hoare.wait(?\textit{bb}$\rightarrow$cons,~?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 26}\'\>~~~~\textit{result}~=~\textit{bb}$\rightarrow$buf[\textit{bb}$\rightarrow$head]\\

\>{\tiny 27}\'\>~~~~\textit{bb}$\rightarrow$head~=~(\textit{bb}$\rightarrow$head~\%~\textit{bb}$\rightarrow$size)~+~1\\

\>{\tiny 28}\'\>~~~~\textit{bb}$\rightarrow$count~--=~1\\

\>{\tiny 29}\'\>~~~~hoare.signal(?\textit{bb}$\rightarrow$prod,~?\textit{bb}$\rightarrow$mon)\\

\>{\tiny 30}\'\>~~~~hoare.exit(?\textit{bb}$\rightarrow$mon)