### 2021-02-02 book/sources/PetersonMethod.tex

\>{\tiny 1}\'\>\texttt{\textbf{def}}~P\_enter(\textit{pm},~\textit{pid}):\\

\>{\tiny 2}\'\>~~~~\textit{pm}$\rightarrow$flags[\textit{pid}]~=~\texttt{\textbf{True}}\\

\>{\tiny 3}\'\>~~~~\textit{pm}$\rightarrow$turn~=~1~--~\textit{pid}\\

\>{\tiny 4}\'\>~~~~\texttt{\textbf{await}}~(\texttt{\textbf{not}}~\textit{pm}$\rightarrow$flags[1~--~\textit{pid}])~\texttt{\textbf{or}}~(\textit{pm}$\rightarrow$turn~==~\textit{pid})\\

\>{\tiny 5}\'\>\\

\>{\tiny 6}\'\>\texttt{\textbf{def}}~P\_exit(\textit{pm},~\textit{pid}):\\

\>{\tiny 7}\'\>~~~~\textit{pm}$\rightarrow$flags[\textit{pid}]~=~\texttt{\textbf{False}}\\

\>{\tiny 8}\'\>\\

\>{\tiny 9}\'\>\texttt{\textbf{def}}~P\_mutex():\\

\>{\tiny 10}\'\>~~~~\textit{result}~=~\{~.turn:~\texttt{\textbf{choose}}(\{0,~1\}),~.flags:~[~\texttt{\textbf{False}},~\texttt{\textbf{False}}~]~\}\\

\>{\tiny 11}\'\>\\

\>{\tiny 12}\'\>\emph{\#\#\#\# The code above can go into its own Harmony module \#\#\#\#}\\

\>{\tiny 13}\'\>\\

\>{\tiny 14}\'\>\textit{mutex}~=~P\_mutex()\\

\>{\tiny 15}\'\>\\

\>{\tiny 16}\'\>\texttt{\textbf{def}}~thread(\textit{self}):\\

\>{\tiny 17}\'\>~~~~\texttt{\textbf{while}}~\texttt{\textbf{choose}}(\{~\texttt{\textbf{False}},~\texttt{\textbf{True}}~\}):\\

\>{\tiny 18}\'\>~~~~~~~~P\_enter(?\textit{mutex},~\textit{self})\\

\>{\tiny 19}\'\>~~~~~~~~@cs:~\texttt{\textbf{assert}}~\texttt{\textbf{atLabel}}(cs)~==~\{~(thread,~\textit{self}):~1~\}\\

\>{\tiny 20}\'\>~~~~~~~~P\_exit(?\textit{mutex},~\textit{self})\\

\>{\tiny 21}\'\>\\

\>{\tiny 22}\'\>\texttt{\textbf{spawn}}~thread(0)\\

\>{\tiny 23}\'\>\texttt{\textbf{spawn}}~thread(1)