### 2021-02-02 book/sources/RWsbs.tex

\>{\tiny 1}\'\>\texttt{\textbf{from}}~synch~\texttt{\textbf{import}}~BinSema,~acquire,~release\\

\>{\tiny 2}\'\>\\

\>{\tiny 3}\'\>\texttt{\textbf{def}}~RWlock():\\

\>{\tiny 4}\'\>~~~~\textit{result}~=~\{\\

\>{\tiny 5}\'\>~~~~~~~~~~~~.nreaders:~0,~.nwriters:~0,~.mutex:~BinSema(\texttt{\textbf{False}}),\\

\>{\tiny 6}\'\>~~~~~~~~~~~~.r\_gate:~\{~.sema:~BinSema(\texttt{\textbf{True}}),~.count:~0~\},\\

\>{\tiny 7}\'\>~~~~~~~~~~~~.w\_gate:~\{~.sema:~BinSema(\texttt{\textbf{True}}),~.count:~0~\}\\

\>{\tiny 8}\'\>~~~~~~~~\}\\

\>{\tiny 9}\'\>\\

\>{\tiny 10}\'\>\texttt{\textbf{def}}~release\_one(\textit{rw}):\\

\>{\tiny 11}\'\>~~~~\texttt{\textbf{if}}~(\textit{rw}$\rightarrow$nwriters~==~0)~\texttt{\textbf{and}}~(\textit{rw}$\rightarrow$r\_gate.count~$>$~0):\\

\>{\tiny 12}\'\>~~~~~~~~release(?\textit{rw}$\rightarrow$r\_gate.sema)\\

\>{\tiny 13}\'\>~~~~\texttt{\textbf{elif}}~((\textit{rw}$\rightarrow$nreaders~+~\textit{rw}$\rightarrow$nwriters)~==~0)~\texttt{\textbf{and}}~(\textit{rw}$\rightarrow$w\_gate.count~$>$~0):\\

\>{\tiny 14}\'\>~~~~~~~~release(?\textit{rw}$\rightarrow$w\_gate.sema)\\

\>{\tiny 15}\'\>~~~~\texttt{\textbf{else}}:\\

\>{\tiny 16}\'\>~~~~~~~~release(?\textit{rw}$\rightarrow$mutex)\\

\>{\tiny 17}\'\>\\

\>{\tiny 18}\'\>\texttt{\textbf{def}}~read\_acquire(\textit{rw}):\\

\>{\tiny 19}\'\>~~~~acquire(?\textit{rw}$\rightarrow$mutex)\\

\>{\tiny 20}\'\>~~~~\texttt{\textbf{if}}~\textit{rw}$\rightarrow$nwriters~$>$~0:\\

\>{\tiny 21}\'\>~~~~~~~~\textit{rw}$\rightarrow$r\_gate.count~+=~1;~release\_one(\textit{rw})\\

\>{\tiny 22}\'\>~~~~~~~~acquire(?\textit{rw}$\rightarrow$r\_gate.sema);~\textit{rw}$\rightarrow$r\_gate.count~--=~1\\

\>{\tiny 23}\'\>~~~~\textit{rw}$\rightarrow$nreaders~+=~1\\

\>{\tiny 24}\'\>~~~~release\_one(\textit{rw})\\

\>{\tiny 25}\'\>\\

\>{\tiny 26}\'\>\texttt{\textbf{def}}~read\_release(\textit{rw}):\\

\>{\tiny 27}\'\>~~~~acquire(?\textit{rw}$\rightarrow$mutex);~\textit{rw}$\rightarrow$nreaders~--=~1;~release\_one(\textit{rw})\\

\>{\tiny 28}\'\>\\

\>{\tiny 29}\'\>\texttt{\textbf{def}}~write\_acquire(\textit{rw}):\\

\>{\tiny 30}\'\>~~~~acquire(?\textit{rw}$\rightarrow$mutex)\\

\>{\tiny 31}\'\>~~~~\texttt{\textbf{if}}~(\textit{rw}$\rightarrow$nreaders~+~\textit{rw}$\rightarrow$nwriters)~$>$~0:\\

\>{\tiny 32}\'\>~~~~~~~~\textit{rw}$\rightarrow$w\_gate.count~+=~1;~release\_one(\textit{rw})\\

\>{\tiny 33}\'\>~~~~~~~~acquire(?\textit{rw}$\rightarrow$w\_gate.sema);~\textit{rw}$\rightarrow$w\_gate.count~--=~1\\

\>{\tiny 34}\'\>~~~~\textit{rw}$\rightarrow$nwriters~+=~1\\

\>{\tiny 35}\'\>~~~~release\_one(\textit{rw})\\

\>{\tiny 36}\'\>\\

\>{\tiny 37}\'\>\texttt{\textbf{def}}~write\_release(\textit{rw}):\\

\>{\tiny 38}\'\>~~~~acquire(?\textit{rw}$\rightarrow$mutex);~\textit{rw}$\rightarrow$nwriters~--=~1;~release\_one(\textit{rw})