### 2021-02-02 book/sources/qsort.tex

\>{\tiny 1}\'\>\texttt{\textbf{def}}~Qsort(\textit{arr}):\\

\>{\tiny 2}\'\>~~~~\textit{result}~=~\{~.arr:~arr,~.todo:~\{~(0,~\texttt{\textbf{len}}(arr)~--~1)~\}~\}\\

\>{\tiny 3}\'\>\\

\>{\tiny 4}\'\>\texttt{\textbf{def}}~swap(\textit{p},~\textit{q}):~~~~~~~~~~~~~~~\emph{\# swap !p and !q}\\

\>{\tiny 5}\'\>~~~~!\textit{p},~!\textit{q}~=~!\textit{q},~!\textit{p};~\\

\>{\tiny 6}\'\>\\

\>{\tiny 7}\'\>\texttt{\textbf{def}}~partition(\textit{qs},~\textit{lo},~\textit{hi}):\\

\>{\tiny 8}\'\>~~~~\textit{result}~=~\textit{lo}\\

\>{\tiny 9}\'\>~~~~\texttt{\textbf{for}}~\textit{i}~\texttt{\textbf{in}}~\{\textit{lo}..\textit{hi}~--~1\}:\\

\>{\tiny 10}\'\>~~~~~~~~\texttt{\textbf{if}}~\textit{qs}$\rightarrow$arr[\textit{i}]~$<=$~\textit{qs}$\rightarrow$arr[\textit{hi}]:\\

\>{\tiny 11}\'\>~~~~~~~~~~~~swap(?\textit{qs}$\rightarrow$arr[\textit{result}],~?\textit{qs}$\rightarrow$arr[\textit{i}])\\

\>{\tiny 12}\'\>~~~~~~~~~~~~\textit{result}~+=~1\\

\>{\tiny 13}\'\>~~~~swap(?\textit{qs}$\rightarrow$arr[\textit{result}],~?\textit{qs}$\rightarrow$arr[\textit{hi}]);~\\

\>{\tiny 14}\'\>\\

\>{\tiny 15}\'\>\texttt{\textbf{def}}~sortrange(\textit{qs},~\textit{range}):\\

\>{\tiny 16}\'\>~~~~\texttt{\textbf{let}}~\textit{lo},~\textit{hi}~=~\textit{range}~\texttt{\textbf{let}}~\textit{pivot}~=~partition(\textit{qs},~\textit{lo},~\textit{hi}):\\

\>{\tiny 17}\'\>~~~~~~~~\texttt{\textbf{if}}~(\textit{pivot}~--~1)~$>$~\textit{lo}:\\

\>{\tiny 18}\'\>~~~~~~~~~~~~\textit{qs}$\rightarrow$todo~$\vert$=~\{~(\textit{lo},~\textit{pivot}~--~1)~\}\\

\>{\tiny 19}\'\>~~~~~~~~\texttt{\textbf{if}}~(\textit{pivot}~+~1)~$<$~\textit{hi}:\\

\>{\tiny 20}\'\>~~~~~~~~~~~~\textit{qs}$\rightarrow$todo~$\vert$=~\{~(\textit{pivot}~+~1,~\textit{hi})~\}\\

\>{\tiny 21}\'\>\\

\>{\tiny 22}\'\>\texttt{\textbf{def}}~sort(\textit{qs}):\\

\>{\tiny 23}\'\>~~~~\texttt{\textbf{while}}~\textit{qs}$\rightarrow$todo~!=~\{\}:\\

\>{\tiny 24}\'\>~~~~~~~~\texttt{\textbf{let}}~\textit{range}~=~\texttt{\textbf{choose}}(\textit{qs}$\rightarrow$todo):\\

\>{\tiny 25}\'\>~~~~~~~~~~~~\textit{qs}$\rightarrow$todo~--=~\{~\textit{range}~\}\\

\>{\tiny 26}\'\>~~~~~~~~~~~~sortrange(\textit{qs},~\textit{range})\\

\>{\tiny 27}\'\>~~~~\textit{result}~=~\textit{qs}$\rightarrow$arr