### 2021-02-02 book/sources/qsorttest.tex

\>{\tiny 1}\'\>\texttt{\textbf{import}}~qsort,~\texttt{\textbf{bag}}\\

\>{\tiny 2}\'\>\\

\>{\tiny 3}\'\>\texttt{\textbf{const}}~NITEMS~=~4\\

\>{\tiny 4}\'\>\\

\>{\tiny 5}\'\>\textit{a}~=~[~\texttt{\textbf{choose}}(\{1..NITEMS\})~\texttt{\textbf{for}}~\textit{i}~\texttt{\textbf{in}}~\{1..\texttt{\textbf{choose}}(\{1..NITEMS\})\}~]\\

\>{\tiny 6}\'\>\textit{testqs}~=~qsort.Qsort(\textit{a})\\

\>{\tiny 7}\'\>\textit{sa}~=~qsort.sort(?\textit{testqs})\\

\>{\tiny 8}\'\>\texttt{\textbf{assert}}~\texttt{\textbf{all}}(\textit{sa}[\textit{i}~--~1]~$<=$~\textit{sa}[\textit{i}]~\texttt{\textbf{for}}~\textit{i}~\texttt{\textbf{in}}~\{1..\texttt{\textbf{len}}(\textit{sa})--1\})~\emph{\# sorted?}\\

\>{\tiny 9}\'\>\texttt{\textbf{assert}}~\texttt{\textbf{bag}}.fromList(\textit{a})~==~\texttt{\textbf{bag}}.fromList(\textit{sa});~\emph{\# is it a permutation?}