### 2021-02-02 book/sources/triangle.tex

\>{\tiny 1}\'\>\texttt{\textbf{const}}~N~=~10\\

\>{\tiny 2}\'\>\\

\>{\tiny 3}\'\>\texttt{\textbf{def}}~triangle(\textit{n}):~~~\emph{\# computes the n'th triangle number}\\

\>{\tiny 4}\'\>~~~~\textit{result}~=~0\\

\>{\tiny 5}\'\>~~~~\texttt{\textbf{for}}~\textit{i}~\texttt{\textbf{in}}~\{1..\textit{n}\}:~~~~~~\emph{\# for each integer from 1 to n inclusive}\\

\>{\tiny 6}\'\>~~~~~~~~\textit{result}~+=~\textit{i}~~~~~~\emph{\# add i to result}\\

\>{\tiny 7}\'\>\\

\>{\tiny 8}\'\>\textit{x}~=~\texttt{\textbf{choose}}(\{0..N\})~~~~~~~\emph{\# select an x between 0 and N inclusive}\\

\>{\tiny 9}\'\>\texttt{\textbf{assert}}~triangle(\textit{x})~==~((\textit{x}~*~(\textit{x}~+~1))~/~2)